Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-27 paper.tex
Fill out author information and run spell-check.
Fixes in figure 1
Add environments to figure 2.
reuse environment definition in PL&LC
edits in section 2

%% For double-blind review submission, w/o CCS and ACM Reference (max 
%%submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max 
%%submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\stxrule}[3]{\text{#2} & #1 & ::= & #3\\}
\newcommand{\stxrulecont}[1]{& & | & #1 \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\hifunrule}[3]{\highlight{#1} &\highlight{=}& \highlight{#2} & 
\highlight{#3}\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\comprulel}[4]{#1 & \fatsemi^l & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\judgeType}[2]{#1 : #2}
\newcommand{\judgeTypeFT}[3]{#1 : #2 \Longrightarrow #3} % FT = From To
\newcommand{\lazyUD}{Lazy\;UD}
\newcommand{\lazyD}{Lazy\;D}

@@

\author{Andre Kuhlenschmidt}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
  \stxrule{T}{Types}{
    \star \mid{}
    P
  }
  \stxrule{P}{Pre-types}{
    \POOunit \mid
    \POOfun{T}{T} \mid
    \POOprod{T}{T} \mid
    \POOsum{T}{T}
  }
  \stxrule{e}{Terms}{
    \eOOvar{x} \mid{}
    \eOOsole{} \mid{}
    \eOOlam{T}{T}{x}{e} \mid
    \eOOapp{e}{e} \mid
    \eOOcons{T}{T}{e}{e} \mid
    \eOOcar{e} \mid
    \eOOcdr{e}
  }
  \stxrulecont{
    \eOOinl{T}{T}{e} \mid
    \eOOinl{T}{T}{e} \mid
    \eOOcase{e}{e}{e} \mid
    \eOOcast{e}{T}{l}{T} \mid
    \eOOblame{l}
  }
  \end{array}
  \]

  Consistency
  \fbox{$T_1 \sim T_2$}
  \begin{gather*}
  \inference{}{
    \star \sim \star
  } \quad
  \inference{}{
    \star \sim P
  } \quad
  \inference{}{
    P \sim \star
  } \\[1ex]
  \inference{}{
    \iota \sim \iota
  } \quad
  \inference{
    S_1 \sim S_2 &
    T_1 \sim T_2
  }{
    S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
  } \quad
  \inference{
    S_1 \sim S_2 &
    T_1 \sim T_2
  }{
    S_1 \times T_1 \sim S_2 \times T_2
  } \quad
  \inference{
    S_1 \sim S_2 &
    T_1 \sim T_2
  }{
    S_1 \plus T_1 \sim S_2 \plus T_2
  }
  \end{gather*}

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a casted function, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly takes out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split a casted injection, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

@@

\begin{figure}
  Machine state and other runtime data structures
  \[
  \begin{array}{lrcl}
  \stxrule{v}{Values}{
    \vOOtt{} \mid
    \vOOfun{\mathcal{E}}{x}{e} \mid
    \vOOcons{v}{v} \mid
    \vOOinl{v} \mid
    \vOOinr{v} \mid   
    \vOOcast{v}{c}
  }
  \stxrule{c}{Casts}{
    \cOOcast{T}{l}{T}
  }
  \stxrule{I}{Injectable types (\lazyD)}{
    P
  }
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOunit \mid
    \POOfun{\star}{\star} \mid
    \star \times \star \mid
    \star + \star
  }
  \stxrule{o}{Observations}{
    \oOOinj \mid
    \oOOsole \mid
    \oOOfun \mid
    \oOOcons \mid
    \oOOinl \mid
    \oOOinr \mid
    \oOOblame{l}
  }
  \stxrule{r}{Cast results}{
    \rOOsucc{v} \mid
    \rOOfail{l}
  }
  \stxrule{s}{States}{
    \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
    \sOOreturn{v}{\kappa} \mid{}
    \sOOhalt{o}
  }
  \stxrule{\kappa}{Continuations}{
    \kOOmt \mid
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
    \kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa}
  }
  \stxrulecont{ 
    \kOOcdr{\kappa}\mid
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa}\mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa}
  }
  \stxrulecont{
    \kOOcaseIII{v}{v}{\kappa} \mid
    \kOOcast{c}{\kappa}
  }
  \end{array}
  \]

        Shallow-consistency
  \fbox{$T \smile T$}
  \begin{gather*}
  \inference{}{
    \star \smile \star
  } \quad
  \inference{}{
    \star \smile P
  } \quad
  \inference{}{
    P \smile \star
  } \\
  \inference{}{
    \iota \smile \iota
  } \quad
  \inference{}{
    T_{11} \rightarrow T_{12} \smile T_{21} \rightarrow T_{22}
  } \quad
  \inference{}{
    T_{11} \times T_{12} \smile T_{21} \times T_{22}
  } \quad
  \inference{}{
  T_{11} \plus T_1 \smile S_2 \plus T_2
  }
  \end{gather*}

@@

  \caption{\textit{cast} and its auxilliary functions for Lazy UD.}
  \label{fig:HC-UD-cast}
\end{figure}

>>>>

%% For double-blind review submission, w/o CCS and ACM Reference (max

%%submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max 
%%submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

@@

\newcommand{\GTLC}{\texttt{GTLC+}}
\newcommand{\figref}[1]{Fig.~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\stxinrule}[3]{\text{#2} & #1 & \in & #3\\}
\newcommand{\stxrule}[3]{\text{#2} & #1 & ::= & #3\\}
\newcommand{\stxrulecont}[1]{& & | & #1 \\}
\newcommand{\funrule}[3]{#1 &=& #2 & #3\\}
\newcommand{\hifunrule}[3]{\highlight{#1} &\highlight{=}& \highlight{#2} & 
\highlight{#3}\\}
\newcommand{\comprule}[4]{#1 & \fatsemi^\ell & #2 & = & #3 & #4 \\}
\newcommand{\comprulel}[4]{#1 & \fatsemi^l & #2 & = & #3 & #4 \\}
\newcommand{\plus}[0]{+}
\newcommand{\judgetype}[3]{#1 \vdash #2 : #3}
\newcommand{\judgeType}[2]{#1 : #2}
\newcommand{\judgeTypeFT}[3]{#1 : #2 \Longrightarrow #3} % FT = From To
\newcommand{\lazyUD}{Lazy\;UD}
\newcommand{\lazyD}{Lazy\;D}

@@

\author{Andre Kuhlenschmidt}
\email{akuhlens@iu.edu}         %% \email is recommended
\affiliation{
  %\position{Position2b}
  \department{Computer Science Department}             %% \department is recommended
  \institution{Indiana University}           %% \institution is required
  %\streetaddress{Street3b Address2b}
  %\city{City2b}
  %\state{State2b}
  %\postcode{Post-Code2b}
  \country{United States}                   %% \country is recommended
}

@@

\begin{figure}
  Syntax
  \[
  \begin{array}{lrcl}
  \stxrule{T, S}{Types}{
    \star \mid{}
    P
  }
  \stxrule{P}{Pre-types}{
    \POOunit \mid
    \POOfun{T}{T} \mid
    \POOprod{T}{T} \mid
    \POOsum{T}{T}
  }
  \stxrule{e}{Terms}{
    \eOOvar{x} \mid{}
    \eOOsole{} \mid{}
    \eOOlam{T}{T}{x}{e} \mid
    \eOOapp{e}{e} \mid
    \eOOcons{T}{T}{e}{e} \mid
    \eOOcar{e} \mid
    \eOOcdr{e}
  }
  \stxrulecont{
    \eOOinl{T}{T}{e} \mid
    \eOOinl{T}{T}{e} \mid
    \eOOcase{e}{e}{e} \mid
    \eOOcast{e}{T}{l}{T} \mid
    \eOOblame{l}
  }
  \end{array}
  \]

  Consistency
  \fbox{$T_1 \sim T_2$}
  \begin{gather*}
  \inference{}{
    \star \sim \star
  } \quad
  \inference{}{
    \star \sim P
  } \quad
  \inference{}{
    P \sim \star
  } \\[1ex]
  \inference{}{
    \POOunit \sim \POOunit
  } \quad
  \inference{
    S_1 \sim S_2 &
    T_1 \sim T_2
  }{
    S_1 \rightarrow T_1 \sim S_2 \rightarrow T_2
  } \quad
  \inference{
    S_1 \sim S_2 &
    T_1 \sim T_2
  }{
    S_1 \times T_1 \sim S_2 \times T_2
  } \quad
  \inference{
    S_1 \sim S_2 &
    T_1 \sim T_2
  }{
    S_1 \plus T_1 \sim S_2 \plus T_2
  }
  \end{gather*}

@@

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a casted function, the machine first casts $v_1$, the operand, then
applies $v_2$ (the underlying function) to the casted operand, and finally cast
the return value.
%
To take out the first (resp. second) part of a casted pair, the
machine firstly takes out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split a casted injection, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

@@

\begin{figure}
  Machine state and other runtime data structures
  \[
  \begin{array}{lrcl}
  \stxrule{\mathcal{E}}{Environments}{
  	\text{a partial function}\;\{\langle x , v \rangle , \dots \}
%  	\text{partial functions that map  variables to values}
  }
  \stxrule{v}{Values}{
    \vOOtt{} \mid
    \vOOfun{\mathcal{E}}{x}{e} \mid
    \vOOcons{v}{v} \mid
    \vOOinl{v} \mid
    \vOOinr{v} \mid   
    \vOOcast{v}{c}
  }
  \stxrule{c}{Casts}{
    \cOOcast{T}{l}{T}
  }
  \stxrule{I}{Injectable types (\lazyD)}{
    P
  }
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOunit \mid
    \POOfun{\star}{\star} \mid
    \star \times \star \mid
    \star + \star
  }
  \stxrule{o}{Observations}{
    \oOOinj \mid
    \oOOsole \mid
    \oOOfun \mid
    \oOOcons \mid
    \oOOinl \mid
    \oOOinr \mid
    \oOOblame{l}
  }
  \stxrule{r}{Cast results}{
    \rOOsucc{v} \mid
    \rOOfail{l}
  }
  \stxrule{s}{States}{
    \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
    \sOOreturn{v}{\kappa} \mid{}
    \sOOhalt{o}
  }
  \stxrule{\kappa}{Continuations}{
    \kOOmt \mid
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
    \kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa}
  }
  \stxrulecont{ 
    \kOOcdr{\kappa}\mid
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa}\mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa}
  }
  \stxrulecont{
    \kOOcaseIII{v}{v}{\kappa} \mid
    \kOOcast{c}{\kappa}
  }
  \end{array}
  \]

        Shallow-consistency
  \fbox{$T \smile T$}
  \begin{gather*}
  \inference{}{
    \star \smile \star
  } \quad
  \inference{}{
    \star \smile P
  } \quad
  \inference{}{
    P \smile \star
  } \\
  \inference{}{
    \POOunit \smile \POOunit
  } \quad
  \inference{}{
    T_{11} \rightarrow T_{12} \smile T_{21} \rightarrow T_{22}
  } \quad
  \inference{}{
    T_{11} \times T_{12} \smile T_{21} \times T_{22}
  } \quad
  \inference{}{
  T_{11} \plus T_1 \smile S_2 \plus T_2
  }
  \end{gather*}

@@

  \caption{\textit{cast} and its auxiliary functions for Lazy UD.}
  \label{fig:HC-UD-cast}
\end{figure}
